Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-26 latex/article.tex
ok

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\txtblue{
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface,
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
need to be done.}
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

\vspace{0.2cm}

@@

\vspace{0.2cm}

@@

\vspace{0.2cm} \txtblue{In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose an
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real-time and present our results.}

@@

In~\cite{DGCI2013}, \comJeremy{the 2D digital curvature estimator $\CurvH{R}$
and the 3D digital mean curvature estimator $\MeanCurvH{R}$ are based on the
digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):}
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), \emph{digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \vspace{-0.05cm}
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
\comJeremy{Estimators of both 2D curvature and 3D mean curvature on digital
objects with volumetric integral information are proved to be multigrid
convergent~\cite{DGCI2013}.}
%
For short, when the digital object becomes finer and finer, \ie when the
digitization step $h$ tends to zero, the estimated quantity on
$\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
\txtblue{ associated one on $\Bd{\Shape}$ in $O\left(h^{\frac{1}{3}}\right)$ for
convex shapes with at least $C^3$-boundary and bounded curvature (setting
$R\EqDef kh^\frac{1}{3}$ for some $k\in\R$).}
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

In~\cite{CVIU2014}, authors have also defined 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z \EqDef \DSh$ of a shape $\Shape \subset \R^3$,
  \emph{3D digital principal curvature estimators} are defined for any point $\vx
  \in \R^3$ as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    \PrincCurvH{1}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5 R} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5 R} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues
  of the covariance matrix of \comJeremy{$\Ball{R/h}{\vx/h} \cap Z$.}
  % the intersection between the digital shape
  % $\DigShape$ and a ball of digital radius $R/h$ centered on digital point
  % $\vx/h$.
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, $1$
and $2$ (see Equation~20 of~\cite{CVIU2014} for more details).
% ($\DMom{000}{h}$), of order $1$ ($\DMom{100}{h}$, $\DMom{010}{h}$ and
% $\DMom{001}{h}$) and of order $2$ ($\DMom{200}{h}$, $\DMom{020}{h}$,
% $\DMom{002}{h}$, $\DMom{110}{h}$, $\DMom{011}{h}$ and $\DMom{101}{h}$, see
% Equation~20 of \cite{CVIU2014} for more details).
\txtblue{
These estimators are proven convergent in $O\left(h^{\frac{1}{3}}\right)$ when setting the ball
radius $h$ in $R = kh^\frac{1}{3}$, where $k$ is a constant related to the
maximal curvature of the shape, on convex shapes with at least $C^3$-boundary
and bounded curvature~\cite{CVIU2014}.}
%More formally, it leads to:
%%
%\begin{align}
%  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
%  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
%\end{align}

Additionally, eigenvectors associated to $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ of the covariance matrix are principal
curvature direction estimators $\PrincDirH{1}{R}$ and
$\PrincDirH{2}{R}$. The smallest eigenvector corresponds to the normal
direction $\NormalDirH{R}$ at $\vx$. Convergence results can be found
in~\cite{ChapterIICurvature}.
%%% Changer avec le chapitre de Jaco à un moment

@@

\txtblue{\sout{We describe here two different visualization pipeline on GPU that will
be evaluated in the next section.}}

\paragraph{Hierarchical Decomposition.}
\txtblue{
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$.
%(Fig.~\ref{fig:approx}-$middle$)
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$).
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities.
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$.
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx}-$middle$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section~\ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}}

@@

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.45\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.45\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_normales}}

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\txtblue{
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface,
and is naturally related to the amount allowed perturbations on input
data. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
need to be done.}
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

@@

\txtblue{In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose an
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real-time and present our results.}

@@

In~\cite{DGCI2013}, \comJeremy{authors define a  2D digital curvature estimator $\CurvH{R}$
and a 3D digital mean curvature estimator $\MeanCurvH{R}$  based on the
digital volume estimator $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (area estimator $\AreaC(Y,h)$ in dimension 2):}
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), \emph{digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \vspace{-0.05cm}
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
\comJeremy{Such curvature estimators have multigrid convergence
  properties~\cite{DGCI2013}:} For short, when
the digital object becomes finer and finer, \ie when the digitization
step $h$ tends to zero, the estimated quantity on
$\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to
the \txtblue{ associated one on $\Bd{\Shape}$ in
  $O\left(h^{\frac{1}{3}}\right)$ for convex shapes with at least
  $C^3$-boundary and bounded curvature (setting $R\EqDef
  kh^\frac{1}{3}$ for some $k\in\R$).}
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

In~\cite{CVIU2014}, authors have also defined 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z \EqDef \DSh$ of a shape $\Shape \subset \R^3$,
  \emph{3D digital principal curvature estimators} are defined for any point $\vx
  \in \R^3$ as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    \PrincCurvH{1}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5 R} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5 R} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues
  of the covariance matrix of \comJeremy{$\Ball{R/h}{\vx/h} \cap Z$.}
  % the intersection between the digital shape
  % $\DigShape$ and a ball of digital radius $R/h$ centered on digital point
  % $\vx/h$.
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, $1$
and $2$ (see Equation~20 of~\cite{CVIU2014} for more details).
% ($\DMom{000}{h}$), of order $1$ ($\DMom{100}{h}$, $\DMom{010}{h}$ and
% $\DMom{001}{h}$) and of order $2$ ($\DMom{200}{h}$, $\DMom{020}{h}$,
% $\DMom{002}{h}$, $\DMom{110}{h}$, $\DMom{011}{h}$ and $\DMom{101}{h}$, see
% Equation~20 of \cite{CVIU2014} for more details).
\txtblue{
These estimators are proven convergent in $O\left(h^{\frac{1}{3}}\right)$ when setting the ball
radius $h$ in $R = kh^\frac{1}{3}$, where $k$ is a constant related to the
maximal curvature of the shape, on convex shapes with at least $C^3$-boundary
and bounded curvature~\cite{CVIU2014}.}
%More formally, it leads to:
%%
%\begin{align}
%  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
%  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
%\end{align}
Additionally, eigenvectors associated to $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ of the covariance matrix are principal
curvature direction estimators $\PrincDirH{1}{R}$ and
$\PrincDirH{2}{R}$. The smallest eigenvector corresponds to the normal
direction $\NormalDirH{R}$ at $\vx$. Convergence results can be found
in~\cite{ChapterIICurvature}.
%%% Changer avec le chapitre de Jaco à un moment

@@

%\txtblue{\sout{We describe here two different visualization pipeline on GPU that will
 %   be evaluated in the next section.}}

\paragraph{Hierarchical Decomposition.}
\txtblue{
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$.
%(Fig.~\ref{fig:approx}-$middle$)
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$).
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities.
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$.
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx}-$middle$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section~\ref{sec:discussion}).}
%\sout{ Still, this method is not fully
%exploiting the GPU parallelism.  The hardware optimizes the
%parallelism only when the micro-program described in the shader has a
%predictive execution flow. Hence, implementing this recursive
%algorithm (or its de-recursified version) requires a lot of branches
%(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}}

@@

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.4\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.4\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_normales}}
